---
title: "Centaurea Project"
author: "DG, IM, BS"
date: "29/03/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the Rmarkdown file for the Centaurea project containing the code required to replicate the analysis and the plots.

We begin by loading the required packages for the rest of the script.

```{r library, message = FALSE}
library(ggeffects)
library(gridExtra)
library(cowplot)
library(lme4)
library(MuMIn)
library(sjPlot)
library(tidyverse)

```

Next, we load in our dataset.

```{r data, message = FALSE}
data <- read_tsv("donnees_centauree.csv")

data <- data %>%
    mutate(light = PAR/reference) %>%
    rename("species" = "espece",
           "population" = "pop")

data$species <- factor(data$species, levels = c("mac", "cor"))

head(data)
```

We will now split the dataset into two part. The first part will be used to model the probability of germination, and the second part will be used to model the rosette size growth.

```{r datasplit, message=FALSE}

didnotgerm <- data %>%
  filter(is.na(Date_de_germination)) %>%
  mutate(germination = 0)

didgerm <- data %>%
  filter(!is.na(Date_de_germination)) %>%
  mutate(germination = 1)

germ_data <- full_join(didnotgerm, didgerm) %>%
  select(-c(Date_de_germination:reference))

rm(didnotgerm, didgerm)

head(germ_data)

rosette_data <- data %>%
    rename("2005-12-01" = "Taille_Dec_05",
         "2006-02-01" = "Taille_Fev_06",
         "2006-03-01" = "Taille_Mars_06",
         "2006-06-01" = "Taille_Juin_06",
         "2006-09-01" = "Taille_Sept_06") %>%
  gather(key = "Date", 
         value = "Rosette_size", 
         "2005-12-01":"2006-09-01") %>%
  select(-c(traitement:reference)) %>%
  group_by(Plante) %>%
  filter(Rosette_size == max(Rosette_size),
         Rosette_size > 0) %>%
  distinct(Plante, .keep_all = TRUE)

rosette_data$Date_de_germination <- as.Date(rosette_data$Date_de_germination, format = "%d/%m/%Y")
rosette_data$Date <- as.Date(rosette_data$Date, format = "%Y-%m-%d")

rosette_data <- rosette_data %>%
  mutate(time_after_germ = difftime(Date, Date_de_germination, units = "days")) %>%
  filter(time_after_germ > 0)

rosette_data$time_after_germ <- as.numeric(rosette_data$time_after_germ)
  
rm(data)

head(rosette_data)
  
```

Now, we will build and plot the germination model.

```{r germ_mod, fig.align='center', fig.width=5, fig.height=5}

germ_mod1 <- glmer(germination~light*species+(1+species|population), data = germ_data, family = "binomial")
germ_mod2 <- glmer(germination~light*species+(1|population), data = germ_data, family = "binomial")
germ_mod3 <- glm(germination~light*species, data = germ_data, family = "binomial")

anova(germ_mod1, germ_mod2) # (1+species|population) not sig better
anova(germ_mod2, germ_mod3) # however including (1|population) is supported

summary(germ_mod2) # all terms are sig

germ_mod4 <- update(germ_mod2, .~. -light:species)
anova(germ_mod2, germ_mod4)

AICc(germ_mod1, germ_mod2, germ_mod3, germ_mod4)
logLik(germ_mod4)

germ_mod <- germ_mod2

rm(germ_mod1, germ_mod2, germ_mod3, germ_mod4)

germ_pred <- ggpredict(germ_mod, terms = c("light [all]", "species"), type = "fixed") %>%
  rename("species" = "group")

germ_pred$species <- recode_factor(germ_pred$species, mac = "C. maculosa", cor = "C. corymbosa")

germplot <- ggplot(data = germ_data,
       aes(x = light, y = germination)) +
  geom_point(data = germ_data %>%
               filter(species == "mac"),
             aes(x = light, y = germination),
             size = 3, alpha = 0.1, colour = "#F8766D") +
  geom_point(data = germ_data %>%
               filter(species == "cor"),
             aes(x = light, y = germination),
             size = 3, alpha = 0.1, colour = "#00BFC4") +
  geom_line(data = germ_pred,
            aes(y = predicted, x = x, colour = species),
            size = 1.5) +
  geom_ribbon(data = germ_pred, 
              aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high, colour = species, fill = species),
              alpha = 0.1) +
  labs(y = "P(Germination)",
       x = "Light (received PAR / reference PAR)") +
  scale_x_continuous(breaks = c(0,0.25,0.5,0.75,1)) +
  theme_cowplot() +
  theme(legend.position = c(0.65,0.25),
        legend.title = element_blank())

ggsave(filename = "Figure2.png",
       plot = germplot,
       dpi = 600,
       width = 5,
       height = 5,
       type="cairo-png"
       )

rm(germ_pred, germ_mod, germplot)
```

And finally, we build and plot the model for rosette size.

```{r rosette_mod}

rosette_mod1 <- lmer(Rosette_size ~ light*species*Cotyledons*time_after_germ+(1+species|population),
                     data = rosette_data)

rosette_mod2 <- lmer(Rosette_size ~ light*species*Cotyledons*time_after_germ+(1|population),
                     data = rosette_data)

rosette_mod3 <- lm(Rosette_size~light*species*Cotyledons*time_after_germ,
                   data = rosette_data)

anova(rosette_mod1, rosette_mod2) # no difference in L.ratio, AIC support simpler model
anova(rosette_mod2, rosette_mod3) # no difference in L.ratio again, and AIC would marginally support simpler model
# going with simpler = better if both explain the data just as well, we will move on with a lm

rosette_mod4 <- lm(Rosette_size~light*species*Cotyledons*time_after_germ, data = rosette_data)
summary(rosette_mod4)

rosette_mod5 <- update(rosette_mod4, .~. -light:species:Cotyledons:time_after_germ) # try remove 4 way int
anova(rosette_mod4, rosette_mod5) # removal is supported by lrt
summary(rosette_mod5)

rosette_mod6 <- update(rosette_mod5, .~. -species:Cotyledons:time_after_germ) # try remove non sig 3 way int
anova(rosette_mod5, rosette_mod6) # removal supported by lrt
summary(rosette_mod6)

rosette_mod7 <- update(rosette_mod6, .~. -light:Cotyledons:time_after_germ) # remove non sig 3 way int
anova(rosette_mod6,rosette_mod7) # removal supported by lrt
summary(rosette_mod7)

rosette_mod8 <- update(rosette_mod7, .~. -light:species:Cotyledons) # remove non sig 3 way int
anova(rosette_mod7,rosette_mod8) # removal supported by lrt
summary(rosette_mod8)

rosette_mod9 <- update(rosette_mod8, .~. -Cotyledons:time_after_germ) # remove non sig int
anova(rosette_mod8,rosette_mod9) # removal supported by lrt
summary(rosette_mod9)

rosette_mod10 <- update(rosette_mod9, .~. -species:time_after_germ)
anova(rosette_mod9,rosette_mod10) # removal supported by lrt
summary(rosette_mod10)

rosette_mod11 <- update(rosette_mod10, .~. -species:Cotyledons) # remove non sig int
anova(rosette_mod10,rosette_mod11) # removal supported by lrt
summary(rosette_mod11)

rosette_mod12 <- update(rosette_mod11, .~. -light:Cotyledons) # remove marg. sig. int
anova(rosette_mod11,rosette_mod12) # removal supported marginally
summary(rosette_mod12) # all terms sig or are included in sig ints

rosette_mod13 <- update(rosette_mod12, .~. -light:species:time_after_germ)
anova(rosette_mod12, rosette_mod13) # supported to be kept in

rosette_mod14 <- update(rosette_mod12, .~. -light:species)
anova(rosette_mod12, rosette_mod14) # supported to be kept in

rosette_mod <- rosette_mod12 

plot_model(rosette_mod, type = "diag")

mod_list <- paste0("rosette_mod",1:14)
Akaike <- NULL
LL <- NULL

for ( i in 1:length(mod_list)){
Akaike[i] = AICc(get(mod_list[i]),k = 2)
LL[i] = logLik(get(mod_list[i]))
}

mod_sum <- data.frame(mod_list, Akaike, LL)
mod_sum

# recheck random effect structure with min. ad. mod

re1 <- lmer(Rosette_size ~ light + species + Cotyledons + time_after_germ + 
    light:species + light:time_after_germ + light:species:time_after_germ+(1+species|population), 
    data = rosette_data)


re2 <- lmer(Rosette_size ~ light + species + Cotyledons + time_after_germ + 
    light:species + light:time_after_germ + light:species:time_after_germ+(1|population), 
    data = rosette_data)

anova(re1, re2)
anova(re2, rosette_mod) # still not worth including



samp_test <- data.frame(matrix(nrow = 8))

byspe <- rosette_data %>%
  group_by(species)

for (i in 1:1000) {
  samp <- sample_n(byspe, 100)
lm_samp <- lm(Rosette_size ~ light + species + Cotyledons + time_after_germ + 
    light:species + light:time_after_germ + light:species:time_after_germ, data = samp)

samp_test <- cbind(samp_test, lm_samp$coefficients)
  
}

samp <- sample_n(byspe, 113)
lm_samp <- lm(Rosette_size ~ light + species + Cotyledons + time_after_germ + 
    light:species + light:time_after_germ + light:species:time_after_germ, data = samp)

samp_test <- cbind(samp_test, lm_samp$coefficients)

colnames(samp_test) <- paste0("Model",1:ncol(samp_test))

samp_test <- samp_test %>%
  select(-Model1)

samples <- as.data.frame(t(samp_test))

summary(samples)

hist(samples$`(Intercept)`)
hist(samples$light)
hist(samples$speciescor)
hist(samples$Cotyledons)
hist(samples$time_after_germ)
hist(samples$`light:speciescor`)
hist(samples$`light:time_after_germ`)
hist(samples$`light:speciescor:time_after_germ`)

rm(list = mod_list)
rm(re1, re2, mod_list, Akaike, i, LL, samp, samp_test, samples)

```

```{r figures_rosette, fig.align='center',fig.width=12, fig.height=8, message=FALSE, warning=FALSE}
rosette_pred1 <- ggpredict(rosette_mod, terms = c("light", "species"),type = "fixed") %>%
  rename("species" = "group")

rosette_pred2 <- ggpredict(rosette_mod, terms = c("Cotyledons", "species"),type = "fixed") %>%
  rename("species" = "group")

rosette_pred3 <- ggpredict(rosette_mod, terms = c("time_after_germ", "species"),type = "fixed") %>%
  rename("species" = "group")

rosette_pred1$species <- recode_factor(rosette_pred1$species, mac = "C. maculosa", cor = "C. corymbosa")
rosette_pred2$species <- recode_factor(rosette_pred2$species, mac = "C. maculosa", cor = "C. corymbosa")
rosette_pred3$species <- recode_factor(rosette_pred3$species, mac = "C. maculosa", cor = "C. corymbosa")

A <- ggplot(data = rosette_data,
             aes(x = light, y = Rosette_size)) +
  geom_point(data = rosette_data %>%
               filter(species == "mac"),
             aes(x = light, y = Rosette_size),
             size = 4, alpha = 0.3, colour = "#F8766D") +
  geom_point(data = rosette_data %>%
               filter(species == "cor"),
             aes(x = light, y = Rosette_size),
             size = 4, alpha = 0.3, colour = "#00BFC4") +
  geom_line(data = rosette_pred1,
            aes(y = predicted, x = x, colour = species),
            size = 1.5) +
  geom_ribbon(data = rosette_pred1,
              aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high, colour = species, fill = species),
              alpha = 0.1) +
  labs(y = "Maximum rosette size (mm)",
       x = "Light (received PAR / reference PAR)",
       title = "A") +
  scale_x_continuous(expand = c(0, 0), limits = c(0,1.09)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0,150)) +
  theme_cowplot()+
  theme(legend.position = "none")

B <- ggplot(data = rosette_data,
             aes(x = Cotyledons, y = Rosette_size)) +
  geom_point(data = rosette_data %>%
               filter(species == "mac"),
             aes(x = Cotyledons, y = Rosette_size),
             size = 3, alpha = 0.3, colour = "#F8766D") +
  geom_point(data = rosette_data %>%
               filter(species == "cor"),
             aes(x = Cotyledons, y = Rosette_size),
             size = 3, alpha = 0.3, colour = "#00BFC4") +
  geom_line(data = rosette_pred2,
            aes(y = predicted, x = x, colour = species),
            size = 1.5) +
  geom_ribbon(data = rosette_pred2,
              aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high, colour = species, fill = species),
              alpha = 0.1) +
  labs(y = "Maximum rosette size (mm)",
       x = "Cotyledon size (mm)",
       title = "B") +
  scale_x_continuous(expand = c(0, 0), limits = c(0,26)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0,150)) +
  theme_cowplot() +
  theme(axis.title.y=element_blank(),
       # axis.text.y=element_blank(),
        legend.position = "none",
        legend.title = element_blank())

C <- ggplot(data = rosette_data,
             aes(x = time_after_germ, y = Rosette_size)) +
  geom_point(data = rosette_data %>%
               filter(species == "mac"),
             aes(x = time_after_germ, y = Rosette_size),
             size = 3, alpha = 0.3, colour = "#F8766D") +
  geom_point(data = rosette_data %>%
               filter(species == "cor"),
             aes(x = time_after_germ, y = Rosette_size),
             size = 3, alpha = 0.3, colour = "#00BFC4") +
  geom_line(data = rosette_pred3,
            aes(y = predicted, x = x, colour = species),
            size = 1.5) +
  geom_ribbon(data = rosette_pred3,
              aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high, colour = species, fill = species),
              alpha = 0.1) +
  labs(y = "Maximum rosette size (mm)",
       x = "Time after germination (days)",
       title = "C") +
  scale_x_continuous(expand = c(0, 0), limits = c(0,325)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0,150)) +
  theme_cowplot() +
  theme(axis.title.y=element_blank(),
        legend.position = c(0.05,0.9),
        legend.title = element_blank())

layout <- rbind(c(1,1,2),
                c(1,1,3))

ABC <- grid.arrange(A,B,C, layout_matrix=layout)

ggsave(filename="Figure1.png",
       plot = ABC,
       dpi = 600,
       width = 12,
       height = 8,
       type="cairo-png")

rm(A, B, C, ABC, rosette_mod, rosette_pred1, rosette_pred2, rosette_pred3)
```

